On a quotient topology of the partition lattice with forbidden block sizes

### 2019-05-16 paper2/paper.tex
mistakes in paper2 fixed
extract for all, mistake in paper2

Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. We call the vertices of $B$ the \emph{blue} vertices and the vertices of $R$ the \emph{red} vertices. Let $\Delta^R$ be the subtrisp of $\Delta$ that contains all simplices $\sigma\in{\cal F}(\Delta^R)$ with $V(\sigma)\subset R$. For $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ let $m(\sigma)$ denote the smallest blue vertex of $\sigma$.
\begin{df}
A map $\varphi:B\longrightarrow R$ is called a \emph{closure map} if for any $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ either $m(\sigma)\in V(\sigma)$ or $m(\sigma)$ can be uniquely inserted into $\sigma$, i.e. there exists a unique simplex $\tau\in{\cal F}(\Delta)$ with $m(\sigma)\in V(\tau)$ and $\sigma$ is a subsimplex of $\tau$ with $\dim\tau=\dim\sigma+1$.
\end{df}
\begin{thm}[Kozlov \cite{clmap}]
\label{clthm}
Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. Assume we have a closure map $\varphi:B\longrightarrow R$, then there exists an acyclic matching on ${\cal F}(\Delta)$ such that a simplex $\sigma\in{\cal F}(\Delta)$ is critical if and only if $\sigma\in{\cal F}(\Delta^R)$. In particular $\Delta$ collapses onto its subtrisp $\Delta^R$.
\end{thm}
\begin{rem}
\label{coll}
If a trisp $\Delta$ collapses onto a subtrisp $\Delta'$, then $\Delta$ is homotopy equivalent to $\Delta'$. If $\Delta'$ consists of one vertex, we call $\Delta$ \emph{collapsible}.
\end{rem}
\section{The main result}
Let $n\geq 3$ and $2\leq k<n$ be a fixed natural numbers throughout this section.

The simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ are represented by sequences of partitions, where the partitions refine each other. Such a sequence can be considered as a leveled forest, where each level corresponds to a partition. Each node on a particular level corresponds to a block of the corresponding partition in the sequence. Now we modify such a forest as follows. We replace each node, which is a subset of $[n]$, by its cardinality and put a mark onto the number if this subset contains $1$. It is easy to see that the simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ can be indexed with such forests. The vertices can be indexed with number partitions, which we may write as $v_0\oplus v_1+\dots+v_r$, of $n$ that distinguish the first number, i.e. $\oplus$ is non-commutative. The number on the left side of $\oplus$, that is $v_0$, corresponds to the block that contains $1$. We call this number the \emph{number on the left side}. All other numbers are on the right side. The boundary operators are obtained by deleting entire levels from forests and reconnecting vertices transitively through the deleted levels.
\begin{figure}[ht]
\centering
\input{forest.pdf_t} 
\caption{A simplex of $\Delta(\Pi_{6,3})/S_1\times S_5$. One number is distinguished on each level.}
\end{figure}

@@

Now assume $n\equiv 0,1\mod k$. Each vertex of $\Delta_{n,k}$ can be written as
\[(ak+b)\oplus k^c+1^{n-ak-b-ck}\]
with $b\in\{0,1\}$, $a,b,c\in\Z$, $a,b,c\geq 0$, $a+b>0$, $a+c>0$, $ak+b+ck\leq n$ and $ak+b<n$ such that $(a,b,c)$ is unique. The two last inequalities can be rewritten as $0<a+c\leq\frac nk-\frac bk$ and $a<\frac nk-\frac bk$. Now we would like to eliminate $\frac bk$ from these inequalities. For $n\equiv 0\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac nk-b\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac nk
\end{eqnarray*}
and for $n\equiv 1\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac {n-1}k\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac {n-1}k+1-b
\end{eqnarray*}
Hence in the case $n\equiv 0\mod k$ the set of tuples $(a,b,c)$ only depends on $\frac nk$ and in the case $n\equiv 1\mod k$ this set only depends on $\frac {n-1}k$. In particular the number of vertices of $\Delta_{n,k}$ only depends on these numbers.
\begin{lem}
\label{trispiso}
Let $n_i\geq3$ and $2\leq k_i<n_i$ for $i=1,2$. Assume $n_i\equiv 0\mod k_i$ or $n_i\equiv 1\mod k_i$, and assume $\frac{n_1}{k_1}=\frac{n_2}{k_2}$ or $\frac{n_1-1}{k_1}=\frac{n_2-1}{k_2}$, respectively. Then the trisps $\Delta_{n_1,k_1}$ and $\Delta_{n_2,k_2}$ are isomorphic, in particular homotopy equivalent.
\end{lem}
\begin{proof}
We obtain the canonical bijection
\begin{eqnarray*}
\varphi:V(\Delta_{n_1,k_1})&\longrightarrow&V(\Delta_{n_2,k_2})\\
(ak_1+b)\oplus k_1^c+1^{n_1-ak_1-b-ck_1}&\longmapsto&(ak_2+b)\oplus k_2^c+1^{n_2-ak_2-b-ck_2}
\end{eqnarray*}
We also obtain a bijection between ${\cal F}(\Delta_{n_1,k_1})$ and ${\cal F}(\Delta_{n_2,k_2})$ as follows. Let $\sigma\in{\cal F}(\Delta_{n_1,k_1})$ be a simplex, which is a forest. First, we remove all $1$'s from the tree which are blocks on the right side. Second, on the right side we replace each $k_1$ by $k_2$, on the left side we replace $ak_1+b$ by $ak_2+b$. The last step is: We add $1$'s to each level such that the sum over all numbers is $n_2$ and such that they become children of those blocks which do not have enough children. The last step is well-defined. It is easy to see that $\varphi$ is a trisp isomorphism.
\begin{figure}[ht]
\centering
\input{mapping.pdf_t} 
\caption{Example for the trisp isomorphism}
\end{figure}
\end{proof}
\begin{prop}
\label{mainresult}
Let $n\geq3$ and $2\leq k<n$, then the topological space $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is
\begin{enumerate}
\item collapsible for $n\not\equiv 0,1\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2n}k-3$ for $n\equiv 0\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2(n-1)}k-2$ for $n\equiv 1\mod k$.
\end{enumerate}
\end{prop}
\begin{proof}
Since $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is homotopy equivalent to $\Delta_{n,k}$ by Lemma \ref{erste_closure}, we may consider $\Delta_{n,k}$.

@@

We consider the cochain complex of $\Delta$ at the $l$th position, see \cite[Section 3.1]{hatcher}:
\[
\xymatrix{
\dots&\hom(C_{l+1},\Z)\ar[l]&\hom(C_l,\Z)\ar[l]_{\partial^*}&\hom(C_{l-1},\Z)\ar[l]_{\partial^{**}}&\dots\ar[l]
}
\]
We give is a short description of $\partial^*$. Let $\tau$ be simplex of dimension $l+1$.
\[
(\partial^*f)(\tau)=\sum_{i=1}^{l+2}(-1)^{i+1}f(B_i(\tau))
\]
$H^l(\Delta)$ is defined as $H^l(\Delta):=\ker\partial^*/\im\partial^{**}$.

>>>>

Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. We call the vertices of $B$ the \emph{blue} vertices and the vertices of $R$ the \emph{red} vertices. Let $\Delta^R$ be the subtrisp of $\Delta$ that contains all simplices $\sigma\in{\cal F}(\Delta^R)$ with $V(\sigma)\subset R$. For $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ let $m(\sigma)$ denote the smallest blue vertex of $\sigma$.
\begin{df}
A map $\varphi:B\longrightarrow R$ is called a \emph{closure map} if for any $\sigma\in{\cal F}(\Delta)\setminus{\cal F}(\Delta^R)$ either $\varphi(m(\sigma))\in V(\sigma)$ or $\varphi(m(\sigma))$ can be uniquely inserted into $\sigma$, i.e. there exists a unique simplex $\tau\in{\cal F}(\Delta)$ with $\varphi(m(\sigma))\in V(\tau)$ and $\sigma$ is a subsimplex of $\tau$ with $\dim\tau=\dim\sigma+1$.
\end{df}
\begin{thm}[Kozlov \cite{clmap}]
\label{clthm}
Let $\Delta$ be a regular trisp whose set of vertices is divided into two disjoint parts $B\cup R$. Assume we have a closure map $\varphi:B\longrightarrow R$, then there exists an acyclic matching on ${\cal F}(\Delta)$ such that a simplex $\sigma\in{\cal F}(\Delta)$ is critical if and only if $\sigma\in{\cal F}(\Delta^R)$. In particular $\Delta$ collapses onto its subtrisp $\Delta^R$.
\end{thm}
\begin{rem}
\label{coll}
If a trisp $\Delta$ collapses onto a subtrisp $\Delta'$, then $\Delta$ is homotopy equivalent to $\Delta'$. If $\Delta'$ consists of one vertex, we call $\Delta$ \emph{collapsible}.
\end{rem}
\section{The main result}
Let $n\geq 3$ and $2\leq k<n$ be a fixed natural numbers throughout this section.

The simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ are represented by sequences of partitions, where the partitions refine each other. Such a sequence can be considered as a leveled forest, where each level corresponds to a partition. Each node on a particular level corresponds to a block of the corresponding partition in the sequence. Now we modify such a forest as follows. We replace each node, which is a subset of $[n]$, by its cardinality and put a mark onto the number if this subset contains $1$. It is easy to see that the simplices of $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ can be indexed with such forests. The vertices can be indexed with number partitions, which we may write as $v_0\oplus v_1+\dots+v_r$, of $n$ that distinguish the first number, i.e. $\oplus$ is non-commutative. The number on the left side of $\oplus$, that is $v_0$, corresponds to the block that contains $1$. We call this number the \emph{number on the left side}. All other numbers are on the right side. The boundary operators are obtained by deleting entire levels from forests and reconnecting vertices transitively through the deleted levels.
\begin{figure}[ht]
\centering
\input{forest.pdf_t} 
\caption{A simplex of $\Delta(\Pi_{6,2})/S_1\times S_5$. One number is distinguished on each level.}
\end{figure}

@@

Now assume $n\equiv 0,1\mod k$. Each vertex of $\Delta_{n,k}$ can be written as
\[(ak+b)\oplus k^c+1^{n-ak-b-ck}\]
with $b\in\{0,1\}$, $a,b,c\in\Z$, $a,b,c\geq 0$, $a+b>0$, $a+c>0$, $ak+b+ck\leq n$ and $ak+b<n$ such that $(a,b,c)$ is unique. The two last inequalities can be rewritten as $0<a+c\leq\frac nk-\frac bk$ and $a<\frac nk-\frac bk$. Now we would like to eliminate $\frac bk$ from these inequalities. For $n\equiv 0\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac nk-b\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac nk
\end{eqnarray*}
and for $n\equiv 1\mod k$ we have
\begin{eqnarray*}
a+c\leq\frac nk-\frac bk&\Leftrightarrow&a+c\leq\frac {n-1}k\\
a<\frac nk-\frac bk&\Leftrightarrow&a<\frac {n-1}k+1-b
\end{eqnarray*}
Hence in the case $n\equiv 0\mod k$ the set of tuples $(a,b,c)$ only depends on $\frac nk$ and in the case $n\equiv 1\mod k$ this set only depends on $\frac {n-1}k$. In particular the number of vertices of $\Delta_{n,k}$ only depends on these numbers.
\begin{lem}
\label{trispiso}
Let $n_i\geq3$ and $2\leq k_i<n_i$ for $i=1,2$. Assume $n_i\equiv 0\mod k_i$ or $n_i\equiv 1\mod k_i$, and assume $\frac{n_1}{k_1}=\frac{n_2}{k_2}$ or $\frac{n_1-1}{k_1}=\frac{n_2-1}{k_2}$, respectively. Then the trisps $\Delta_{n_1,k_1}$ and $\Delta_{n_2,k_2}$ are isomorphic, in particular homotopy equivalent.
\end{lem}
\begin{proof}
We obtain the canonical bijection
\begin{eqnarray*}
\varphi:V(\Delta_{n_1,k_1})&\longrightarrow&V(\Delta_{n_2,k_2})\\
(ak_1+b)\oplus k_1^c+1^{n_1-ak_1-b-ck_1}&\longmapsto&(ak_2+b)\oplus k_2^c+1^{n_2-ak_2-b-ck_2}
\end{eqnarray*}
We also obtain a bijection between ${\cal F}(\Delta_{n_1,k_1})$ and ${\cal F}(\Delta_{n_2,k_2})$ as follows. Let $\sigma\in{\cal F}(\Delta_{n_1,k_1})$ be a simplex, which is a forest. First, we remove all $1$'s from the tree which are blocks on the right side. Second, on the right side we replace each $k_1$ by $k_2$, on the left side we replace $ak_1+b$ by $ak_2+b$. The last step is: We add $1$'s to each level such that the sum over all numbers is $n_2$ and such that they become children of those blocks which do not have enough children. The last step is well-defined. It is easy to see that $\varphi$ is a trisp isomorphism.
\begin{figure}[ht]
\centering
\input{mapping_paper2.pdf_t}
\caption{Example for the trisp isomorphism}
\end{figure}
\end{proof}
\begin{prop}
\label{mainresult}
Let $n\geq3$ and $2\leq k<n$, then the topological space $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is
\begin{enumerate}
\item collapsible for $n\not\equiv 0,1\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2n}k-3$ for $n\equiv 0\mod k$.
\item homotopy equivalent to a sphere of dimension $\frac{2(n-1)}k-2$ for $n\equiv 1\mod k$.
\end{enumerate}
\end{prop}
\begin{proof}
Since $\Delta(\Pi_{n,k})/S_1\times S_{n-1}$ is homotopy equivalent to $\Delta_{n,k}$ by Lemma \ref{erste_closure}, we may consider $\Delta_{n,k}$.

@@

We consider the cochain complex of $\Delta$ at the $l$th position, see \cite[Section 3.1]{hatcher}:
\[
\xymatrix{
\dots&\hom(C_{l+1},\Z)\ar[l]&\hom(C_l,\Z)\ar[l]_{\partial^*}&\hom(C_{l-1},\Z)\ar[l]_{\partial^{**}}&\dots\ar[l]
}
\]
We give a short description of $\partial^*$. Let $\tau$ be simplex of dimension $l+1$.
\[
(\partial^*f)(\tau)=\sum_{i=1}^{l+2}(-1)^{i+1}f(B_i(\tau))
\]
$H^l(\Delta)$ is defined as $H^l(\Delta):=\ker\partial^*/\im\partial^{**}$.
